---
title: "coNet: model and analyze networks using co-occurrence data"
output: "md_document"
bibliography: "inst/REFERENCES.bib"
---

* Network models are a useful representation of systems [@Woodward2010].
* Because it is not always possible to directly observe relationships
  (i.e. ecological interactions) spatial or temporal patterns of
  co-occurrence can be used to render models of interaction networks
  [@Araujo2011].
* The *conetto* implements an adaptation of the method developed by
  Araujo et al. 2011 expanded with the application of Bayesian
  probability to generate network models and provides several
  functions useful for comparative analysis of sets of co-occurrence
  based networks. 

# Install

Currently, the package is still in beta but it can be installed via
the *devtools* package.

```{r eval = FALSE}
install.packages("devtools")
devtools::install_github("ECGen/coNet")
```

# Example

## Generate co-occurrence data

The basic usage of the *connetto* package is to generate network
models. To do this, we need a matrix of co-occurrence values, which
will be passed into the `coNet` function:

```{r}
A <- c(1, 1, 0, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1)
B <- c(1, 1, 0, 1, 1, 1, 1, 1, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1)
C <- c(1, 0, 1, 0, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1)
D <- c(0, 1, 1, 1, 1, 1, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0)
M <- data.frame(A, B, C, D)
coNet(M)
```

The `coNet` function returns of square matrix with the same number of
rows and columns equal to the number of columns in the co-occurrence
matrix. There are a couple of features that are useful to note here:

* Notice that the matrix is not symetric, that is, the top and bottom
  halves are not equal. This is because the dependency of two
  variables is based on both their independent occurrences and their
  co-occurrences.
* Also, you will notice that some of the values in the network are
  zero and while some are positive there are also negative
  values. This is because `coNet` conducts a test that compares the
  observed co-occurrences to a theoretical "null" co-occurrence value
  that is analytically determined. If the observed co-occurrences are
  lower, inside or greater than the "null" predicted interval, then
  the resulting conditional . The null interval test can be adjusted
  using the *ci.p* argument (DEFAULT = 95).
* 


# References



