---
title: "coNet: model and analyze networks using co-occurrence data"
output: "md_document"
bibliography: "inst/REFERENCES.bib"
---

* Network models are a useful representation of systems [@Woodward2010].
* Because it is not always possible to directly observe relationships
  (i.e. ecological interactions) spatial or temporal patterns of
  co-occurrence can be used to render models of interaction networks
  [@Araujo2011].
* The *conetto* implements an adaptation of the method developed by
  [@Araujo2011] expanded with the application of Bayesian probability
  to generate network models and provides several functions useful for
  comparative analysis of sets of co-occurrence based networks.

# Install

Currently, the package is still in beta but it can be installed via
the *devtools* package.

```{r eval = FALSE}
install.packages("devtools")
devtools::install_github("ECGen/coNet")
```

# Example


The basic usage of the *connetto* package is to generate network
models. To do this, we need a matrix of co-occurrence values, which
will be passed into the `coNet` function:

```{r}
A <- c(1, 1, 0, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1)
B <- c(1, 1, 0, 1, 1, 1, 1, 1, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1)
C <- c(1, 0, 1, 0, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1)
D <- c(0, 1, 1, 1, 1, 1, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0)
M <- data.frame(A, B, C, D)
coNet(M)
```

The `coNet` function returns of square matrix with the same number of
rows and columns equal to the number of columns in the co-occurrence
matrix. There are a couple of features that are useful to note here:

* First, the matrix is not symetric, that is, the top and bottom
  halves are not equal. This is because the dependency of two
  variables is based on both their independent occurrences and their
  co-occurrences.
* Also, some of the values in the network are zero. This is because
  `coNet` conducts a test that compares the observed co-occurrences to
  a theoretical "null" co-occurrence interval. If the observed
  co-occurrences are within the interval, the conditional probability
  is set equal to the joint probability. The threshold for the null
  interval test can be adjusted using the *ci.p* argument (DEFAULT =
  95).
* Last, some values are negative. This is because the network is
  comprised of the difference between the threshold adjusted
  conditional probabilities minus the observed joint probabilities
  (i.e. *P(A|B) - P(A,B)*). Therefore, if the conditional probability
  is less than the observed joint probability, the resulting
  "relative" value will be negative, while the oposite results in a
  positive value for the network. The "raw", un-relativized, values
  can be returned via the *raw* argument.


## Network comparisons

The `coNet` function was developed for the analysis of replicated
datasets, where multiple networks have been observed. Two other
functions were written to provide a way to analyze these network
sets. 

One function, `meanNet`, will calculate the "mean" or average of a set
of networks. The resul is the cell-wise summation of all matrices
divided by the total number of matrices. This function requires that
the set of networks all have the same dinensionaity and that they are
organized into a list.

```{r}
net.l <- lapply(1:10, function(x) matrix(runif(100), nrow = 10))
meanNet(net.l)
```

Another function, `distNet`, facilitates the calculation of a distance
matrix for a set of networks. Two metrics are provided, Euclidean and
Bray-Curtis. The resulting matrix is the distance among all networks
in the set. Similar to `meanNet`, `distNet` requires that all matrices
have the same dimensionality. The result is a distance, or
dissimilarity in the case of Bray-Curtis, matrix that can now be used
for multivariate analyses, such as ordination.

```{r}
net.d <- distNet(net.l)
net.d
```

# References



