---
title: 'conetto: Create network models in R using co-occurrence data'
tags:
  - R
  - networks
  - co-occurrence
authors:
  - name: Matthew K. Lau
    orcid: 0000-0003-3758-2406
    affiliation: "1, 2"
affiliations:
 - name: Harvard Forest, Harvard University 
   index: 1
 - name: Chinese Academy of Sciences, Shenyang, Liaoning, China 
   index: 2
date: 24 April 2019
bibliography: paper.bib
---


# Introduction

* Network models are one way to represent systems [@Borrett2014].
* It's not always possible to directly observe relationships
  (i.e. ecological interactions) [@Cite].
* In ecology, space can be used as a substitute for time [@Gotelli].
* This assumes that variance of conditions and processes are minimally
  impactful [@Cite].
* This method implements an adaptation of the method developed by
  Araujo et al. 2011 expanded with the application of Bayesian
  probability.

## Quick review of other packages available

```{r}
if (!exists("db")){db <- tools::CRAN_package_db()}
mdf <- data.frame(Package = db[, "Package"], 
                  Title = db[, "Title"], 
                  Description = db[, "Description"])
net.pks <- mdf[grep("network", mdf[, "Description"], ignore.case = TRUE), ]
coo.pks <- net.pks[grepl("co-occur", net.pks[, "Description"], ignore.case = TRUE), ]
as.character(coo.pks[, "Package"])
```

EcoSimR: what methods for network modeling? What Bayesian methods?

Araujo method

# Methods

## Overview

1. Calculate pairwise conditional probabilities $P(S_i|S_j) = \frac{P(S_i,S_j)}{P(S_j)}$
2. Calculate variances $V(S_iS_j) = N * E[P(S_i,S_j)] * (1 -
E[P(S_i,S_j)])$
3. Confidence interval $CI_{95\%} = E[S_iS_j] * Z_{95\%} * \sqrt{V(S_iS_j)}$
4. Edge removal $If CI_{l} \leq [S_iS_j] \leq CI_{u}, then P(S_i|S_j) = P(S_i,S_j)$
5. Re-scaling $D_{ij} = P(S_i|S_j) - P(S_i,S_j)$

## Pairwise Conditional Probabilities

* Uses the conditional probabilities of each species pair, i.e. the
probability of observing one species given an observation of another
species $P(S_i | S_j)$, based on the method developed by
[@Araujo2011]. 
* To calculate conditional probabilities, we quantified the individual
probabilities of species occurrences (i.e. $P(S_i)$ or $P(S_j)$) and
the joint probability of co-occurrences $P(S_i,S_j)$ using the
frequencies of each species and their co-occurrences. We were then
able to calculate the conditional probabilities of each species pair
as $P(S_i|S_j) = \frac{P(S_i,S_j)}{P(S_j)}$, based on the axioms of
probability. 
* This yielded a matrix that could possibly be asymmetric,
i.e. $P(S_i|S_j)$ does not have to be equal to $P(S_j|S_i)$.
* Another property of this matrix is that the diagonal ($S_{ii}$) is
equal to one for all species present and zero for species that were
not observed in any cell.




## Edge Removal and Re-scaling

* We then applied an analytical procedure to remove non-significant
links between species. This procedure determines if the joint
probability of a species pair (i.e. $P(S_i,S_j)$) is different from
zero (Fig.~\ref{fig:conet_method}).
* Here, a confidence interval $CI_{95\%}$ is calculated as as
$CI_{95\%} = E[S_iS_j] * Z_{95\%} * \sqrt{V(S_iS_j)}$, where the
expected frequency of co-occurrences E($S_iS_j$) is the total number
of cells surveyed ($N$) times the independent probabilities of each
species $P(S_i) * P(S_j)$, $Z_{95\%}$ is the Z-score for 95\% from a
Z-distribution and the expected variance of $E(S_iS_j)$ is the total
number of cells times the expected probability of $S_iS_j$ and its
compliment (i.e. $V(S_iS_j) = N * E[P(S_i,S_j)] * (1 -
E[P(S_i,S_j)])$). 
* If the observed number of co-occurrence falls outside of the
confidence interval, the joint probability $P(S_i,S_j)$ is determined
to be equal to the product of the individual probabilities
(i.e. $P(S_i) \dot P(S_j)$), and the conditional probability reduces
to the individual probability of that species $P(S_i)$.
* Therefore, unless the co-occurrence of a species pair falls outside
the confidence interval, the probability that the observation of one
species given the other is no different than simply observing that
species alone. 
* This enables us to remove links from a given network by
re-scaling the resulting conditional probabilities by subtracting the
individual probabilities from the conditional probabilities (i.e. how
different the conditional probability is from the independent
probability), which makes any species with a non-significant
conditional probability zero. 
* The resulting matrix ($\mathbf{D} = D_{ij}$) can be interpreted as
how one species impacts another with zero being no effect and values
less than or greater than zero interpreted as negative and positive
effects, respectively. 
* Here, we will refer to this matrix ($\mathbf{D}$) as an interaction
matrix with the properties that it can be asymmetric (i.e. $P_{ij}$
does not necessarily equal $P_{ji}$), and the diagonal ($P_{ii}$) is
zero (i.e. a species does not influence it's own probability of being
observed).





![Lichen interaction networks were constructed by conducting
  field observations in 1 cm$^2$ cells within a 10 cm$^2$ grid on each
  tree using a checkerboard pattern (grey cells). Thus, a set of $N$
  total cell observations were recorded for each tree with the
  presence or absence of each species recorded for each cell. Applying
  the probability-based network modeling method adapted from
  [@Araujo2011], we calculated the conditional probabilities,
  $P(S_i|S_j)$, for all species pairs and removed (i.e. set equal to
  zero) species pairs whose joint probabilities, $P(S_i S_j)$, were
  not significant using a confidence interval based comparison of
  their observed co-occurrence frequency, $S_iS_j$, to that expected
  due to chance alone, $E[P(S_iS_j)] = P(S_i) P(S_j)$, and
  $P(S_i|S_j)$ reduces to $P(S_i)$, the observed individual
  probability of species $S_i$.](img/lcn_araujo_method.pdf)\begin{figure}[ht]

<!-- \centering -->
<!-- \includegraphics[width=\linewidth]{img/lcn_araujo_method.pdf} -->
<!-- \caption{Lichen interaction networks were constructed by conducting -->
<!--   field observations in 1 cm$^2$ cells within a 10 cm$^2$ grid on each -->
<!--   tree using a checkerboard pattern (grey cells). Thus, a set of $N$ -->
<!--   total cell observations were recorded for each tree with the -->
<!--   presence or absence of each species recorded for each cell. Applying -->
<!--   the probability-based network modeling method adapted from -->
<!--   [@Araujo2011], we calculated the conditional probabilities, -->
<!--   $P(S_i|S_j)$, for all species pairs and removed (i.e. set equal to -->
<!--   zero) species pairs whose joint probabilities, $P(S_i S_j)$, were -->
<!--   not significant using a confidence interval based comparison of -->
<!--   their observed co-occurrence frequency, $S_iS_j$, to that expected -->
<!--   due to chance alone, $E[P(S_iS_j)] = P(S_i) P(S_j)$, and -->
<!--   $P(S_i|S_j)$ reduces to $P(S_i)$, the observed individual -->
<!--   probability of species $S_i$.} -->
<!-- \label{fig:conet_method} -->
<!-- \end{figure} -->



# Results


A simple example using data from Keith et al. 2010.

* Arthopods sampled on individual trees in a common garden
* Import data

```{r }

dat <- read.csv("../data/arth09.csv")
## 
dat <- dat[, -1:-2]
dat[is.na(dat)] <- 0
dat[dat != 0] <- 1
## 
pb.cn <- conetto::coNet(dat, ci.p = 99)
pb.cn <- pb.cn[apply(abs(pb.cn), 1, sum) > 0, apply(abs(pb.cn), 2, sum) > 0]
spp <- rownames(pb.cn)
rownames(pb.cn) <- colnames(pb.cn) <- 1:nrow(pb.cn)
pb.ig <- igraph::graph_from_adjacency_matrix(abs(pb.cn), mode = "directed", weighted = TRUE)
pb.btw <- igraph::betweenness(pb.ig, normalized = TRUE)
names(pb.btw) <- spp
igraph::plot.igraph(pb.ig, vertex.size = (pb.btw*100)^(0.85), 
            vertex.label.color = "black", 
            vertex.boarder.color = "white",
            arrow.size = 0.1)

```


# Discussion

# Citations

# Acknowledgments

# References
